Machine Learning, Building Styles and Property Values

### 2019-09-13 text/manuscript_assa.Rmd
vintage/era replaced by style

---
title: "Machine Learning, Building Vintage and Property Values"
author: "Thies Lindenthal* (University of Cambridge) & Erik B. Johnson (University of Alabama)"
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: Externalities of architecture, machine learning
fig_caption: yes
header-includes:
   - \usepackage{soul}
   - \usepackage{color}
   - \sethlcolor{yellow}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{graphicx}
   - \usepackage{grffile}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{multirow}
   - \doublespacing
   - \usepackage[backend=bibtex, style=authoryear]{biblatex}
fig_width: 6
fig_height: 4
output: 
  pdf_document:
    keep_tex: true
bibliography: references.bib
abstract: "This paper first introduces an algorithm that collects pictures of individual buildings from Google Street View. Second, it trains a deep convolutional neural network (CNN) to classify residential buildings into architectural styles, taking into account spatial dependencies of building vintages. Third, it investigates whether architectural styles influence house prices. For re-sales, the architectural style is a significant determinant of transaction prices while no such effect is found for new buildings. Additionally, we are able to provide guidance on how to detect and overcome some of the limitations of machine learning methods through a large-scale comparison of predictions and expert classifications." 
---

# Introduction
\let\svthefootnote\thefootnote
\let\thefootnote\relax\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for exceptional research assistance. Carolin Schmidt, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp, Pavel Krivenko and participants at the ASSA/AREUEA 2019 annual meeting and seminars at UCL/Bartlett and the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}
\addtocounter{footnote}{-1}\let\thefootnote\svthefootnote
If new developments were only pleasing to the eye, so Britain's *Building Better, Building Beautiful Commission*, nimbyism would cease and housing supply could finally reach the levels demanded by a growing and more affluent population [@economistlineofbeauty2018]. The inaugural chair of this new governmental commission boldly suggests that we should *``build, as our Georgian and Victorian forebears built* [...]. *All objections to new building would slip away in the sheer relief of the public''* [@scruton2018]. Even Prince Charles, in similar spirit, put forward ten principles for urban growth and architecture that emphasize tradition and aesthetics [@princecharles2014]. When governments and princes occupy themselves with beauty, economists surely may take a closer look as well. In this paper, we aim to identify the architectural styles of residential buildings using computer vision techniques and to empirically search for any transaction price premia associated with these styles. If aesthetics were as strong a force in the built environment as claimed, we should indeed find a significant effect on property prices.

This paper assesses the value of architectural aesthetics heads on by classifying the vintage of residential buildings using street level images. Recent work illustrates how street level imagery and machine learning classification is an efficient and powerful combination for measuring previously unobserved characteristics of the urban environment. @Naik2017 describe how neighborhood demographics may impact the physical appearance of neighborhoods. @Gebru2017 use classified vehicle make and model to predict income, race, education, and voting patterns at the precinct level. @Glaeser2016 predict income in New York City. @Naik2016AER create a neighborhood safety based Streetscore which is shown to be highly correlated with neighborhood population density and household income. @Nadai2016 find that greenery and street facing windows contribute to a positive appearance of safety while @Liu2017 evaluate the quality and upkeep of the built environment along Beijing's streets.

@@

In this paper we suggest a method to collect a large number of images of individual UK buildings from Google Street View, classify the depicted buildings using deep convolutional neural networks, combine the derived information with sales price data and, ultimately, estimate marginal prices for the estimated building characteristics. Our work demonstrates that utilizing pretrained convolutional neural networks to detect complex characteristics such as housing vintages with relatively high accuracy comes at low computational costs and is feasible with only modestly-sized training data. Additionally, we illustrate how machine classifications may be sensitive to photo quality by comparing predicted vintages to those made by human experts.

## Building vintage and property values

The aesthetics of different architectural vintages\footnote{We use ``style'',``era'' or ``vintage'' interchangeably in this paper.} make for an easy cocktail party conversation topic -- and they may also influence market prices. Using architectural assessments by human experts, @Buitelaar2017 find a sizable premium of 5\% for new buildings in the Netherlands that refer to traditional styles and a staggering 15\% premium for new buildings that closely follow traditional shapes, facade composition and details. Their study carefully disentangles the architectural style from other unobserved characteristics such as building quality, differences in location or year of construction. These controls are crucial, as earlier work has established that age and vintage variables tend to be highly correlated. @Coulson2008, for instance, suggest a non-parametric estimator and establish a U-shaped age function and distinct price discounts for postwar and contemporary vintages (vis-à-vis more historic vintages). @Francke2017a investigate the depreciation of residential real estate in the Netherlands and decompose land versus structure values singling out the effect of "physical deterioration, functional obsolescence and and vintage effects". They find that buildings from the 1930s carry a strong price premium.

@@

@GlaeserKincaidNaik2018 rely on an different ILSVRC competitor, \emph{Resnet-101} [@He2016]. They reduce the extracted feature vectors to lower dimensionality (1024 to 100 dimensions) based on principal component analysis (PCA). We follow a different strategy and double the dimensionality by including the feature vector of the building's nearest neighbor. Doubling up allows us to model spatial dependencies in building styles, similar to spatially correlated land cover classifications in @Ghimire2010a.

@@

Cambridge's houses can be classified into seven broad eras: \footnote{We are grateful to colleagues from the Architecture Department at the University of Cambridge to provide these general vintage descriptions.}

@@

* In the \emph{Late Victorian} era (c1870s--1901), bay windows became more and more widespread, and increasingly substantial. Stylistic movements such as the Queen Anne revival style contributed richly ornamental details to the formal repertoire employed by designers. Stained glass became more popular. \emph{Edwardian} architecture (1901-1910) tends to be less ornate than late Victorian architecture.

@@

* The \emph{Postwar} (1950-1980) era continued on this path, with an embrace of high-rise as well as low rise housing. Facades vary greatly between brick, tiling, pebbledash and render.

* The cut-off year for our \emph{Contemporary} era to begin is 1980. \emph{Revival} are contemporary buildings trying to emulate historic, mostly Victorian, architecture.

Two final-year architectural students classified a large sub-sample of approximately 25,000 images from our data set of Cambridge houses. This is a much larger sample than needed in our case. In our case, each category requires less than 250 observations to reach saturated training accuracy levels. We greatly exceed this number for the purpose of this paper so that we can compare the out-of-sample convolutional neural network predictions to the ground truth as assigned by the experts. This allows us to examine the power and size of the assignment tests. In addition having both human and machine classification for a large sample of the data allows for robustness checks on the machine comparisons.

We create stratified training samples of 600 buildings from each category but *Georgian*, which is the least common style in Cambridge and for which we can only sample 330 examples. This leaves a sufficiently sized out-of-sample verification data set.

@@

For 100 randomly drawn, stratified data sets, we train two models each: The first is based on the building specific feature vectors (2048 dimensions) only while the second also incorporates the feature vector of the nearest neighbor (combined 4096 dimensions). In cases where trees, fences or large vehicles obstruct the view or where the front of a building cannot be classified with high certainty, the image of the nearest neighbor provides a ``second opinion'' that is spatially correlated with the observation of interest. Finally we classify all buildings not used in training or evaluating the model (truly out-of-sample).

When classifying a building picture, each models returns a vector of vintage-scores, each between 0 to 1, that jointly sum up to 1. We select the vintage with the highest score as the best estimate. By not excluding observations where multiple scores are vying for the top rank we retain as many observations as possible but risk a higher misclassification rate.

@@

We tabulate confusion matrices and evaluate the classification performance in terms of recall, precision and $F_1$-scores. The next step regresses the true vintage observed, captured by binomial variables $D_{TrueVint}$, against a vector of hedonic variables $\mathbf{X}$, vectors of year $\mathbf{Y}$ and neighborhood $\bm{Loc}$ dummy variables and the building's $\mathbf{Vintage}$, estimated both with and without spatial dependencies. The intercept is denoted by $\alpha$ while $\bm{\beta}$, $\bm{\delta}$, $\bm{\gamma}$ and $\bm{\lambda}$ are vectors of regression coefficients, and $\epsilon$ is the error term.

\begin{equation} \label{eq:rev}
  logit( D_{TrueVint,i} ) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Vintage}_i +  \gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

@@

Will the resulting classifiers be biased with respect to property values, picking up correlated but irrelevant cues such as greenery, upkeep, or cars brands instead of building exteriors? Are, for instance, more valuable houses more likely to be classified as the popular *Late Victorian* vintage while starter homes are disproportionally more likely to be classified as e.g. *Early Victorian*? We expect that in cases in which the ensemble provided a high confidence score, building features are clearly detectable on the pictures. For close calls, however, where models within the ensemble cannot agree, other factors might tilt the scale towards an incorrect class. A sizable difference in estimated marginal prices for estimated vintages versus values for ground truth vintages in the subsequent hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

## Building era and property price

Merging the estimated vintage classification with sales data [@LandRegistry2016a] we estimate a hedonic regression equation that establishes marginal prices for the building vintage [similar to @Moorhouse1994; @Asabere1989; @Vandell1989; @Fuerst2011; @Plaut2006], among other characteristics:
\begin{equation} \label{eq:hed}
  ln(Price_i) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Vintage}_i + \eta\mathbf{VintNeigh}_i + \iota\mathbf{Vintage\cdot VintNeigh}_i  +\gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

Here, the natural logarithm of sales prices is explained by a linear combination of hedonic attributes described in vector $\mathbf{X}$, vectors of year $\mathbf{Y}$ and neighborhood $\bm{Loc}$ dummy variables and the building's estimated $\mathbf{Vintage}$ and the prevailing vintage of other buildings in the direct proximity $(\mathbf{VintNeigh})$. The intercept is denoted by $\alpha$ while $\bm{\beta}$, $\bm{\delta}$, $\bm{\eta}$, $\bm{\iota}$, $\bm{\gamma}$ and $\bm{\lambda}$ are vectors of regression coefficients. $\epsilon$ is the IID error term. Heteroscedastic robust standard errors will be reported.

Are buildings from different eras imperfect substitutes catering to multiple groups of households with distinct vintage preferences? @Buitelaar2017 indicate that any premium for an architectural style must stem from either differences in construction prices (which they do not find in their Dutch sample) or from supply constraints, as new construction potentially does not capture the demand for traditional styles. For Cambridge, new supply will inevitably be of either contemporary or revival style as historic vintage buildings, by definition, are not supplied any more. Estimating Eq. \ref{eq:hed} for a subset of newly constructed buildings will show whether construction prices or supply constraints for new homes built according to different architectural styles persist -- if too few vernacular buildings were built, prices should reflect such a shortage.

@@

Overall, we collect pictures of more than 48,000 distinct buildings from the Google Street View API. The subset of 25,000 pictures that has been classified by trained architects into the seven vintages will serve as \emph{ground truth} for our study. Two additional categories identify buildings hidden by *greenery* or uninformative pictures featuring *plain walls*. The classified images are available at the authors' websites.

@@

Almost 16,000 of the 25,000 buildings for which we have the architects' ground truth could be matched to the 23,768 sales transactions that have been recorded for Cambridge between 1995 and 2018. Thus, we observe both the machine *estimated vintage* and architect defined *ground truth* classification for the majority of buildings in the data set.

@@

The agreement between vintage estimates and the corresponding ground truth classifications by experts is reassuringly strong. When relying on feature vectors of the buildings and their closest neighbors, 67 to 84 percent of true classifications are matched with correctly estimated labels (Table \ref{tab:confmat}, Panel A). The recall rates are especially high for older vintages: For *Georgians*, it is 79 percent, *Early Victorian* 81 percent, *Late Victorian/Edwardian* 78 percent, *Interwar* 84 percent, respectively. For more recent *Postwar* buildings, the rate drops to 71 percent, for *Contemporary* to 72 percent and for *Revival* to 67 percent.

A visual inspection of classifications is additionally reassuring. Figure \ref{fig:vint_examples} displays the pictures which carry the highest out-of-sample scores for each vintage. The model is clearly able to differentiate based on small cues, even when only parts of the facade are captured on the picture.

The Herfindahl scores for classifications vary substantially across ground truth vintages (Table \ref{tab:herftab}), indicating that the models within the ensemble, for instance, agree more often on a building being *Georgian* than finding consensus for other vintages. For *Contemporary* or *Revival* vintages, individual classifications differ most. All off-diagonal elements in the lower panel of Table \ref{tab:herftab} are negative, suggesting that, in general, the Herfindahl index is a good predictor of misclassified images.

@@

Table \ref{tab:imgqualsumstats} presents the mean values for the image quality variables. Table \ref{tab:diffimgchar}, Panel A, compares the mean values for views obstructed by trees or vehicles for correctly (on diagonal) and incorrectly classified images. In most cases, the off-diagonal values are positive, showing that more obstacles in the view of sight correlate with a higher likelihood of misclassifications. Similarly, if more of the image area shows a house (Panel B), misclassifications are reduced (negative differences). This effect is stronger for windows, which appear to give valuable cues about a house's vintage (Panel C). Finally, houses that are not in the center of the building tend to be misclassified more frequently (positive differences in Panel D). In sum, bad quality images lead to misclassifications.

@@

The number of misclassifications decreases decidedly when considering observations with high ensemble confidence only: The majority vote based on the most confident two-thirds of predictions only are more in line with ground truth, as higher recall, precision and $F_1$-scores in Table \ref{tab:confmat}, Panel B, show. The improvement is most evident in the precision rates. For the relatively rare categories of *Georgian* and *Revival* about the same number of buildings classified as *Georgian/Revival* in Panel A actually belong to a different ground truth class -- the precision shoots up from 0.50 to 0.70 for *Georgian* and from 0.52 to 0.82 for *Revival* when excluding the not so confident lower third of predictions in Panel B.

@@

Overall, the spatial softmax classification incorporating the nearest neighbor's feature vector outperforms the base classifier: pairwise differences of the $F_1$-scores from the spatial and the base classifier are clearly positive (Fig. \ref{fig:boxplot}). AIC values for 28 binomial models based on Eq. \ref{eq:rev} decrease as hedonic characteristics, location and time dummies are augmented by the base and spatial ML predictor. For all building vintages, the AIC are lowest for the spatial ML predictor. In sum, the spatial classifier performs better for UK vintage data than a classifier based on single buildings only.

@@

The first column in Table \ref{tab:hedreg} presents the estimated regression coefficients for ground truth vintages. The base vintage *Contemporary* is more expensive than almost all other vintages, which show negative coefficients that are significantly different from 0. A clear pecking order appears: *Georgian* and *Revival* buildings demand the highest premium $(+0.04)$,  followed by *Late Victorian/Edwardian* $(+0.01)$, *Contemporary* (base), *Early Victorian* $(-0.13)$,  *Interwar* $(-0.15)$ and *Postwar* $(-0.22)$.

The sample of ground truth images is not necessarily representative of the overall population so we re-estimate the model using predicted vintages for *all* buildings for which we were able to extract usable pictures from Google Street View. Overall, the vintage coefficients become more positive relative to the base, implying that our automatic classification is biased and tends to classify unattractive buildings to the *Contemporary* vintage. Reversely, more attractive buildings are disproportionally often misclassified as *Late Vic./Edw.*, pushing up the corresponding coefficient to $0.11$. This bias vanishes in the third model, which uses only confidently predicted vintages (Column 3).

Models 4--6 follow the same logic as 1--3 but are estimated for *new* buildings only. Basically this contrasts houses which might *look* different from the street, having either contemporary or revival facades, but which are all modern homes at their core. Differences in e.g. materials, floor plans, green space and gardens are minimal. For newly built homes, no effect of *Revival* architecture on price can be found, neither for ground truth classifications nor for high confidence ML estimates. After controlling for location, building characteristics and quality, buyers show no willingness to pay a premium for *Revival* architecture.

Introducing a variable capturing the most prominent style of buildings on the same street and within 100 m ($\mathbf{VintNeigh}$) does not change the pecking order of building vintages much (Table \ref{tab:hedreg}, Column 7). Neighboring styles are compared to the base case of a *Contemporary* house in a *Contemporary* street. The negative coefficients on dissimilar neighboring styles support a finding by @Lindenthal2017 that shows that a harmonious match of a building's shape with its direct environment leads to a price premium -- or a discount in case of shape mismatches. In relative terms, *Postwar* buildings are the least popular neighbors. The difference between *Revival* and *Contemporary* neighbors is not statistically significant. We fail to find evidence in favor of positive externalities from revival architecture. Everything else equal, buildings within ensembles of *Contemporary* generally do not achieve higher transaction prices than buildings in more historic areas (as Table \ref{tab:comb} will confirm later).

Table \ref{tab:hedregint} presents the $\iota$ interaction coefficients (Eq. \ref{eq:hed}) for vintage-neighboring vintage combinations with sufficient numbers of observations (see Table \ref{tab:vintneigh}). The last row suggests that a general claim that *Revival* neighbors exert positive externalities and increase property values relative to *Contemporary* surroundings cannot be confirmed empirically. Again, we find no statistically significant difference.

@@

\emph{Predicted} & \multicolumn{7}{c}{\emph{Ground truth}} \\ 
  \cmidrule(lr){2-8} 
& Georgian & Early Vic. & Late V./Edw. & Interwar & Postwar & Cont. & Revival \\

>>>>

---
title: "Machine Learning, Building Styles and Property Values"
author: "Thies Lindenthal* (University of Cambridge) & Erik B. Johnson (University of Alabama)"
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: Externalities of architecture, machine learning
fig_caption: yes
header-includes:
   - \usepackage{soul}
   - \usepackage{color}
   - \sethlcolor{yellow}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{graphicx}
   - \usepackage{grffile}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{multirow}
   - \doublespacing
   - \usepackage[backend=bibtex, style=authoryear]{biblatex}
fig_width: 6
fig_height: 4
output: 
  pdf_document:
    keep_tex: true
    number_sections: true
bibliography: references.bib
abstract: "This paper first introduces an algorithm that collects pictures of individual buildings from Google Street View. Second, it trains a deep convolutional neural network (CNN) to classify residential buildings into architectural styles, taking into account spatial dependencies of these styles. Third, it investigates whether architectural styles influence house prices. For re-sales, the architectural style is a significant determinant of transaction prices while no such effect is found for new buildings. Additionally, we are able to provide guidance on how to detect and overcome some of the limitations of machine learning methods through a large-scale comparison of predictions and expert classifications." 
---

# Introduction
\let\svthefootnote\thefootnote
\let\thefootnote\relax\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for exceptional research assistance. Carolin Schmidt, Peter Brummund, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp, Pavel Krivenko and participants at the ASSA/AREUEA 2019 annual meeting and seminars at UCL/Bartlett and the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}
\addtocounter{footnote}{-1}\let\thefootnote\svthefootnote
If new developments were only pleasing to the eye, so Britain's *Building Better, Building Beautiful Commission* advising Government, nimbyism would cease and housing supply could finally reach the levels demanded by a growing and more affluent population [@economistlineofbeauty2018]. The inaugural chair of this new commission boldly suggested that if we built *``as our Georgian and Victorian forebears built* [...]. *All objections to new building would slip away in the sheer relief of the public''* [@scruton2018]. Even Prince Charles, in similar spirit, put forward ten principles for urban growth and architecture that emphasize tradition and aesthetics [@princecharles2014]. When governments and princes occupy themselves with beauty, economists surely may take a closer look as well. In this paper, we aim to identify the architectural styles of residential buildings using computer vision techniques and to empirically search for any transaction price premia associated with these styles. If aesthetics were as strong a force in the built environment as claimed, we should indeed find a significant effect on property prices.

This paper assesses the value of architectural aesthetics head on by classifying the exterior styles of residential buildings using street level images. Recent work illustrates how street level imagery and machine learning classification is an efficient and powerful combination for measuring previously unobserved characteristics of the urban environment. @Naik2017 describe how neighborhood demographics may impact the physical appearance of neighborhoods. @Gebru2017 use classified vehicle make and model information to predict income, race, education, and voting patterns at the precinct level. @Glaeser2016 predict income in New York City. @Naik2016AER create a neighborhood safety based Streetscore which is shown to be highly correlated with neighborhood population density and household income. @Nadai2016 find that greenery and street facing windows contribute to a positive appearance of safety while @Liu2017 evaluate the quality and upkeep of the built environment along Beijing's streets.

@@

In this paper we suggest a method to collect a large number of images of individual UK buildings from Google Street View, classify the depicted buildings using deep convolutional neural networks, combine the derived information with sales price data and, ultimately, estimate marginal prices for the estimated building characteristics. Our work demonstrates that utilizing pretrained convolutional neural networks to detect complex characteristics such as housing appearances with relatively high accuracy comes at low computational costs and is feasible with only modestly-sized training data. Additionally, we illustrate how machine classifications may be sensitive to photo quality by comparing machine classifications to those made by human experts.

## Building styles and property values

The aesthetics of different architectural styles make for an easy cocktail party conversation topic -- and they may also influence market prices. Using architectural assessments by human experts, @Buitelaar2017 find a sizable premium of 5\% for new buildings in the Netherlands that refer to traditional styles and a staggering 15\% premium for new buildings that closely follow traditional shapes, facade composition and details. Their study carefully disentangles the architectural style from other unobserved characteristics such as building quality, differences in location, or year of construction. These controls are crucial, as earlier work has established that age and style variables tend to be highly correlated. @Coulson2008, for instance, suggest a non-parametric estimator and establish a U-shaped age function and distinct price discounts for postwar and contemporary styles (vis-à-vis more historic styles). @Francke2017a investigate the depreciation of residential real estate in the Netherlands and decompose land versus structure values singling out the effect of "physical deterioration, functional obsolescence and and vintage effects". They find that buildings from the 1930s carry a strong price premium.

@@

@GlaeserKincaidNaik2018 rely on an different ILSVRC competitor, \emph{Resnet-101} [@He2016]. They reduce the extracted feature vectors to lower dimensionality (1024 to 100 dimensions) based on principal component analysis (PCA). We follow a different strategy and double the dimensionality by including the feature vector of the closest building in the sample, which, in most cases, is the direct neighbor. The feature vectors of neighbors have been collected in exactly the same way as those for other buildings in the sample. Doubling up allows us to model spatial dependencies in building styles, similar to spatially correlated land cover classifications in @Ghimire2010a.

@@

Cambridge's houses can be classified into seven broad styles: \footnote{We are grateful to colleagues from the Architecture Department at the University of Cambridge to provide these general style descriptions.}

@@

* In the \emph{Late Victorian} era (c1870s--1901), bay windows became more and more widespread, and increasingly substantial. Stylistic movements such as the Queen Anne revival style contributed richly ornamental details to the formal repertoire employed by designers. Stained glass became more popular.

* \emph{Edwardian} architecture (1901-1910) tends to be less ornate than late Victorian architecture.

@@

* The \emph{Postwar} (1950-1980) styles continued on this path, with an embrace of high-rise as well as low rise housing. Facades vary greatly between brick, tiling, pebbledash and render.

* The cut-off year for our \emph{Contemporary} era to begin is 1980. \emph{Revival} style-homes are contemporary buildings trying to emulate historic, mostly Victorian, architecture.

Two final-year architectural students classified a large sub-sample of approximately 25,000 images from our data set of Cambridge houses. This is a much larger sample than needed in our case. In our case, each category requires less than 250 observations to reach saturated training accuracy levels. We greatly exceed this number for the purpose of this paper so that we can compare the out-of-sample convolutional neural network predictions to the architects' classifications. This allows us to examine the power and size of the assignment tests. In addition having both human and machine classification for a large sample of the data allows for robustness checks on the machine comparisons.

We create stratified training samples of 600 buildings from each category but *Georgian*, which is the least common style in Cambridge and for which we can only sample 330 examples. This leaves a sufficiently sized out-of-sample verification data set of 21,000 homes.

@@

For 100 randomly drawn data sets, all stratified on style, we train two models each: The first is based on the building specific feature vectors (2048 dimensions) only while the second also incorporates the feature vector of the nearest neighbor (combined 4096 dimensions). In cases where trees, fences or large vehicles obstruct the view or where the front of a building cannot be classified with high certainty, the image of the nearest neighbor provides a ``second opinion'' that is spatially correlated with the observation of interest. Finally we classify all buildings not used in training or evaluating the model (truly out-of-sample).

When classifying a building picture, each models returns a vector of style-scores, each between 0 to 1, that jointly sum up to 1. We select the style with the highest score as the best estimate. By not excluding observations where multiple scores are vying for the top rank we retain as many observations as possible but risk a higher misclassification rate.

@@

We tabulate confusion matrices and evaluate the classification performance in terms of recall, precision and $F_1$-scores. More formally, the next step regresses the style classifications by architects, captured by binomial variables $D_{TrueVint}$, against a vector of hedonic variables $\mathbf{X}$, vectors of year $\mathbf{Y}$ and neighborhood $\bm{Loc}$ dummy variables and the building's $\mathbf{Style}$, estimated both with and without spatial dependencies. The intercept is denoted by $\alpha$ while $\bm{\beta}$, $\bm{\delta}$, $\bm{\gamma}$ and $\bm{\lambda}$ are vectors of regression coefficients, and $\epsilon$ is the error term.

\begin{equation} \label{eq:rev}
  logit( D_{TrueVint,i} ) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Style}_i +  \gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

@@

Will the resulting classifiers be biased with respect to property values, picking up correlated but irrelevant cues such as greenery, upkeep, or cars brands instead of building exteriors? Are, for instance, more valuable houses more likely to be classified as the popular *Late Victorian* style while starter homes are disproportionally more likely to be classified as e.g. *Early Victorian*? We expect that in cases in which the ensemble provided a high confidence score, building features are clearly detectable on the pictures. For close calls, however, where models within the ensemble cannot agree, other factors might tilt the scale towards an incorrect class. A sizable difference in estimated marginal prices for machine's classifications versus the estimates for architects' classifications in the subsequent hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

## Building style and property price

Merging the machine's classification with sales data [@LandRegistry2016a] we estimate a hedonic regression equation that establishes marginal prices for the building style [similar to @Moorhouse1994; @Asabere1989; @Vandell1989; @Fuerst2011; @Plaut2006], among other characteristics:
\begin{equation} \label{eq:hed}
  ln(Price_i) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Style}_i + \eta\mathbf{StyleNeigh}_i + \iota\mathbf{Style\cdot StyleNeigh}_i  +\gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

Here, the natural logarithm of sales prices is explained by a linear combination of hedonic attributes described in vector $\mathbf{X}$, vectors of year $\mathbf{Y}$ and neighborhood $\bm{Loc}$ dummy variables and the building's estimated $\mathbf{Style}$ and the prevailing vintage of other buildings in the direct proximity $(\mathbf{StyleNeigh})$. The intercept is denoted by $\alpha$ while $\bm{\beta}$, $\bm{\delta}$, $\bm{\eta}$, $\bm{\iota}$, $\bm{\gamma}$ and $\bm{\lambda}$ are vectors of regression coefficients. $\epsilon$ is the IID error term. Heteroscedastic robust standard errors will be reported.

Are buildings with different appearances imperfect substitutes catering to multiple groups of households with distinct style preferences? @Buitelaar2017 indicate that any premium for an architectural style must stem from either differences in construction prices (which they do not find in their Dutch sample) or from supply constraints, as new construction potentially does not capture the demand for traditional styles. For Cambridge, new supply will inevitably be of either contemporary or revival style as historic vintages (and their styles), by definition, are not supplied any more. Estimating Eq. \ref{eq:hed} for a subset of newly constructed buildings will show whether construction prices or supply constraints for new homes built according to different architectural styles persist -- if too few vernacular buildings were built, prices should reflect such a shortage.

@@

Overall, we collect pictures of more than 48,000 distinct buildings from the Google Street View API. The subset of 25,000 pictures that has been classified by trained architects into the seven styles will serve as \emph{ground truth} for our study. Two additional categories identify buildings hidden by *greenery* or uninformative pictures featuring *plain walls*. The classified images are available at the authors' websites.

@@

Almost 16,000 of the 25,000 buildings for which we have an architect's classification could be matched to the 23,768 sales transactions that have been recorded for Cambridge between 1995 and 2018. The remaining 9,000 houses have not been sold since 1995 - but the information on their architectural style is still valuable for our measures of neighborhood styles. Thus, we observe both the machine's and the architects' classification for the majority of buildings in the data set.

@@

The agreement between classifications by the machine and by the architects is reassuringly strong. When relying on feature vectors of the buildings and their closest neighbors, 67 to 84 percent of true classifications are matched with correctly estimated labels (Table \ref{tab:confmat}, Panel A). The recall rates are especially high for older styles: For *Georgians*, it is 79 percent, *Early Victorian* 81 percent, *Late Victorian/Edwardian* 78 percent, *Interwar* 84 percent, respectively. For more recent *Postwar* buildings, the rate drops to 71 percent, for *Contemporary* to 72 percent and for *Revival* to 67 percent.

A visual inspection of classifications is additionally reassuring. Figure \ref{fig:vint_examples} displays the pictures which carry the highest out-of-sample scores for each style The model is clearly able to differentiate based on small cues, even when only parts of the facade are captured on the picture.

The Herfindahl scores for classifications vary substantially across styles, as classified by the architects (Table \ref{tab:herftab}), indicating that the models within the ensemble, for instance, agree more often on a building being *Georgian* than finding consensus for other styles. For *Contemporary* or *Revival* styles, individual classifications differ most. All off-diagonal elements in the lower panel of Table \ref{tab:herftab} are negative, suggesting that, in general, the Herfindahl index is a good predictor of misclassified images.

@@

Table \ref{tab:imgqualsumstats} presents the mean values for the image quality variables. Table \ref{tab:diffimgchar}, Panel A, compares the mean values for views obstructed by trees or vehicles for correctly (on diagonal) and incorrectly classified images. In most cases, the off-diagonal values are positive, showing that more obstacles in the view of sight correlate with a higher likelihood of misclassifications. Similarly, if more of the image area shows a house (Panel B), misclassifications are reduced (negative differences). This effect is stronger for windows, which appear to give valuable cues about a house's style (Panel C). Finally, houses that are not in the center of the building tend to be misclassified more frequently (positive differences in Panel D). In sum, bad quality images lead to misclassifications.

@@

The number of misclassifications decreases decidedly when considering observations with high ensemble confidence only: The majority vote based on the most confident two-thirds of machine classifications only are more in line with the architects' classifications, as higher recall, precision and $F_1$-scores in Table \ref{tab:confmat}, Panel B, show. The improvement is most evident in the precision rates. For the relatively rare categories of *Georgian* and *Revival* about the same number of buildings classified as *Georgian/Revival* in Panel A actually belong to a different style, according to the human experts. The precision shoots up from 0.50 to 0.70 for *Georgian* and from 0.52 to 0.82 for *Revival* when excluding the not so confident lower third of predictions in Panel B.

@@

Overall, the spatial softmax classification incorporating the nearest neighbor's feature vector outperforms the base classifier: pairwise differences of the $F_1$-scores from the spatial and the base classifier are clearly positive (Fig. \ref{fig:boxplot}). AIC values for 28 binomial models based on Eq. \ref{eq:rev} decrease as hedonic characteristics, location and time dummies are augmented by the base and spatial ML predictor (Table \ref{tab:aic}). For all building styles, the AIC are lowest for the spatial ML predictor. In sum, the spatial classifier performs better for UK style data than a classifier based on single buildings only.

@@

The first column in Table \ref{tab:hedreg} presents the estimated regression coefficients for the architects' classifications. The base style *Contemporary* is more expensive than almost all other styles, which show negative coefficients that are significantly different from 0. A clear pecking order appears: *Georgian* and *Revival* buildings demand the highest premium $(+0.04)$,  followed by *Late Victorian/Edwardian* $(+0.01)$, *Contemporary* (base), *Early Victorian* $(-0.13)$,  *Interwar* $(-0.15)$ and *Postwar* $(-0.22)$.

The sample images classified by architects is not necessarily representative of the overall population so we re-estimate the model using machine classifications for *all* buildings for which we were able to extract usable pictures from Google Street View. Overall, the style coefficients become more positive relative to the base, implying that our automatic classification is biased and tends to classify unattractive buildings to the *Contemporary* style. Reversely, more attractive buildings are disproportionally often misclassified as *Late Vic./Edw.*, pushing up the corresponding coefficient to $0.11$. This bias vanishes in the third model, which uses only confidently estimated machine styles (Column 3).

Models 4--6 follow the same logic as 1--3 but are estimated for *new* buildings only. Basically this contrasts houses which might *look* different from the street, having either contemporary or revival facades, but which are all modern homes at their core. Differences in e.g. materials, floor plans, green space and gardens are minimal. For newly built homes, no effect of *Revival* architecture on price can be found for high confidence ML estimates, neither for classifications by machines nor by architects. After controlling for location, building characteristics and quality, buyers show no willingness to pay a premium for *Revival* architecture (Columns 4 and 6).

Introducing a variable capturing the most prominent style of buildings on the same street and within 100 m ($\mathbf{VintNeigh}$) does not change the pecking order of building styles much (Table \ref{tab:hedreg}, Column 7). Neighboring styles are compared to the base case of a *Contemporary* house in a *Contemporary* street. The negative coefficients on dissimilar neighboring styles support a finding by @Lindenthal2017 that shows that a harmonious match of a building's shape with its direct environment leads to a price premium -- or a discount in case of shape mismatches. In relative terms, *Postwar* buildings are the least popular neighbors. The difference between *Revival* and *Contemporary* neighbors is not statistically significant. We fail to find evidence in favor of positive externalities from revival architecture. Everything else equal, buildings within ensembles of *Contemporary* generally do not achieve higher transaction prices than buildings in more historic areas (as Table \ref{tab:comb} will confirm later).

Table \ref{tab:hedregint} presents the $\iota$ interaction coefficients (Eq. \ref{eq:hed}) for style-neighboring style combinations with sufficient numbers of observations (see Table \ref{tab:vintneigh}). The last row suggests that a general claim that *Revival* neighbors exert positive externalities and increase property values relative to *Contemporary* surroundings cannot be confirmed empirically. Again, we find no statistically significant difference.

@@

\emph{Machine} & \multicolumn{7}{c}{\emph{Architects}} \\ 
  \cmidrule(lr){2-8} 
& Georgian & Early Vic. & Late V./Edw. & Interwar & Postwar & Cont. & Revival \\
